---
title: "Final Project: Predicting Lifetime Expectancy"
author: "Group 4"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE, warning = FALSE)
```

#Data Loading
```{r}
#load necessary libraries
library(tidyverse) #readr, dplyr

#Visual
library(ggplot2)
library(corrplot)


#Analysis
library(leaps) #regsubsets
library(car) #boxcox
library(caret) #confusion matrix

#load processed dataset
df <- read_csv("df.csv")


#change outcome to categorical
df <- df %>% 
  mutate(Life_Expectancy = case_when(
  Lifetime < 72  ~ 'Below Average',
  Lifetime < 100  ~ 'Above Average')) %>% 
  mutate_at("Life_Expectancy", as.factor)

#delete innecessary variables
df <- df %>% 
  select(-contains("Death")) %>% #not analyzing any of the cause of death factor this time 
  rename("Depression"="Prevalence - Depressive disorders - Sex: Both - Age: Age-standardized (Percent)") %>%
  select(-contains("Prevalence")) %>% #excluding all mental illness term except for depression
  rename("Diet_Animal"="Calories from animal protein (FAO (2017))") %>% 
  rename("Diet_Plant"="Calories from plant protein (FAO (2017))") %>% 
  select(-contains("Calories")) %>% 
  select(-Population,-Continent)

#Change "Above Average" to the target level, since we are interested in longevity
df$Life_Expectancy <- factor(df$Life_Expectancy, levels = c("Below Average","Above Average"))
contrasts(df$Life_Expectancy)
```


#Feature Selection
```{r,message=FALSE,warning=FALSE}


#get the names of crucial features of Sequential, Forward and Backward Selection

#exclude two categorical variable, and year variable, which is predetermined to be included in the data. 
seqrep <- 
  regsubsets(Life_Expectancy~Inventory+Exports+Consumption+Government+Household+Imports, data = df %>% select(-1,-2), 
                     nvmax = 3, method = "seqrep") %>% 
  summary()
seqrep

#forward selection
forward <- 
  regsubsets(Life_Expectancy~Inventory+Exports+Consumption+Government+Household+Imports, data = df %>% select(-1,-2), 
                      nvmax = 3, method = "forward") %>% 
  summary()
forward


#backward selection
backward <- 
  regsubsets(Life_Expectancy~Inventory+Exports+Consumption+Government+Household+Imports, data = df %>% select(-1,-2), 
                      nvmax = 3, method = "backward") %>% 
  summary()

backward
```

#Confusion Matrix
```{r}
df %>% select(Consumption, Education, Diet_Animal, Diet_Plant, Depression, Smoking, Obesity) %>% 
  cor() %>% 
  corrplot()
chisq.test(df$Education,df$Consumption) #testing some high correlation predictors, turns out good! no wrorries
```
 
follow the result above, we delete gdp terms except consumption, also delete diet_animal due to confusion matrix
```{r}
df <- df %>%  
  select(-c(Inventory,Exports,Government,Household,Imports,Diet_Animal))
```

#EDA symbox
```{r}
df %>% ggplot(aes(Consumption))+geom_histogram()
symbox(~Consumption, data=df)
df %>% ggplot(aes(log(Consumption)))+geom_histogram()
#log Consumption

df %>% ggplot(aes(Education))+geom_histogram()
symbox(~Education, data=df)
#keep education the same

df %>% ggplot(aes(Diet_Plant))+geom_histogram()
symbox(~Diet_Plant, data=df)
df %>% ggplot(aes(log(Diet_Plant)))+geom_histogram()
#data is large 1700+ so we are good

df %>% ggplot(aes(Depression))+geom_histogram()
symbox(~Depression, data=df)
df %>% ggplot(aes(log(Depression)))+geom_histogram()
#data is large 1700+ so we are good

df %>% ggplot(aes(Sanitation))+geom_histogram()
symbox(~Sanitation, data=df)
```



```{r}
#view response variable
tab <- data.frame(table(df$Life_Expectancy))
a <- ggplot(tab,aes(x=Var1,y=Freq)) +
  geom_bar(stat="identity", fill="steelblue",width=0.4) +
  theme_bw() + xlab("\nLife Expectancy") + ylab("Frequency\n") +
  ggtitle("Barplot of Life Expectancy Frequency") +
  theme(plot.title = element_text(hjust = 0.5,size=14),
        axis.title.x = element_text(size = 12.5),
        axis.title.y = element_text(size = 12.5)) +
  geom_text(aes(label = Freq), vjust = 1.5, color="white",
             position = position_dodge(.9), size = 4)
a
```

```{r}
library(ggpubr)
c <- ggplot(df,aes(x=Consumption)) +
  geom_histogram(fill="deepskyblue",bins=20)
d <- ggplot(df,aes(x=Diet_Plant)) +
  geom_histogram(fill="cornsilk",bins=20)
e <- ggplot(df,aes(x=Education)) +
  geom_histogram(fill="darkgoldenrod1",bins=20)
f <- ggplot(df,aes(x=Smoking)) +
  geom_histogram(fill="coral1",bins=20)
g <- ggplot(df,aes(x=Depression)) +
  geom_histogram(fill="#999999",bins=20)
h <- ggplot(df,aes(x=Sanitation)) +
  geom_histogram(bins=20,fill="darkolivegreen2")
ggarrange(c,d,e,f,g,h)
```
```{r}
#heatmap for GDP-related
library(reshape2)
corr_mat <- round(cor(df[,4:9]),3)
melted_corr_mat <- melt(corr_mat)
ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  geom_text(aes(Var2, Var1, label = value),
          color = "white", size = 4)
```

```{r}
#heatmap for others
corr_mat <- df %>% 
  select(Consumption, Education,Diet_Plant,Depression, Smoking, Obesity) %>% 
  cor() %>% round(3)
melted_corr_mat <- melt(corr_mat)
ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  geom_text(aes(Var2, Var1, label = value),
          color = "white", size = 4)
```

```{r}
#convert smoking and sanitation
Smoking1 = cut(df$Smoking,breaks=(c(0,13.6,22.6,40)),labels=c("Low","Medium","High"))
Sanitation1 = cut(df$Sanitation,breaks=(c(0,0.95,1)),labels=c("Lack Access","Safe Access"),right = TRUE)
df <- df %>% 
  mutate(Smoking = as.factor(Smoking1)) %>% 
  mutate(Sanitation = as.factor(Sanitation1)) %>% 
  #change order
  select(-Smoking,-Sanitation,-Life_Expectancy,Smoking,Sanitation, Life_Expectancy)
```

```{r}
library(formattable)
formattable(table(df$Smoking,df$Life_Expectancy))
#versus continent
tab1 <- data.frame(table(df[,c("Smoking","Life_Expectancy")]))
b <- ggplot(tab1,aes(x=Smoking,y=Freq,fill=Life_Expectancy)) +
  geom_bar(stat="identity",alpha = 1.5,position = position_dodge()) +
  theme_bw() + ylab("Frequency\n") + scale_fill_brewer(palette="Blues") +
  ggtitle("Barplot of Life Expectancy Frequency v.s Smoking") +
  theme(plot.title = element_text(hjust = 0.5,size=14),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12)) + xlab("\n Smoking") + 
  geom_text(aes(label = Freq), vjust = -0.3,
             position = position_dodge(.9), size = 3.5)
b
```

```{r}
#year
#table(df$Year)
tab2 <- data.frame(table(df[,c("Sanitation","Life_Expectancy")]))
y <- ggplot(tab2,aes(x=Sanitation,y=Freq,fill=Life_Expectancy)) +
  geom_bar(stat="identity",alpha = 1,position = position_dodge()) +
  theme_bw() + ylab("Frequency\n") + 
  scale_fill_manual(values=c("#bae4b3","#238b45")) +
  ggtitle("Barplot of Life Expectancy Frequency v.s Sanitation") +
  theme(plot.title = element_text(hjust = 0.5,size=14),
        axis.title.x = element_text(size = 12.5),
        axis.title.y = element_text(size = 12.5)) + 
  xlab("\n Sanitation Accessibility") +
  geom_text(aes(label = Freq), vjust = -0.3,
             position = position_dodge(.9), size = 4)
y
```









#Model 1
```{r}
mod0 <- glm(Life_Expectancy~log(Consumption)+Education+Diet_Plant+Depression+Obesity+Smoking+Sanitation,df,family="binomial")
summary(mod0)

```

#VIF Model 1
```{r}
vif(glm(Life_Expectancy ~ log(Consumption)+ Education + Diet_Plant + Depression + Obesity, df, family="binomial"))
```

#Final Model
```{r}
mod <- glm(Life_Expectancy~log(Consumption)+Education+Diet_Plant+Obesity+Depression+Smoking+Sanitation+ Smoking*Sanitation,df,family="binomial")
summary(mod)
```

#Effect Plot and Interaction Plot
```{r}
library(effects) 
plot(Effect("Consumption",mod))
plot(Effect("Education",mod))
plot(Effect("Diet_Plant",mod))
plot(Effect("Obesity",mod))
plot(Effect("Depression",mod))
# life <- ifelse(df$Life_Expectancy=="Above Average",1,0)
# Obe <- ifelse(df$Obesity>=49.7,"High","Low")
# Edu <- ifelse(df$Education >8,"High","Low")
# Dep <- ifelse(df$Depression >3.9,"High","Low")
interaction.plot( 
    df$Smoking,#x-axis variable
    df$Sanitation,#variable for line
    df$Lifetime) #y-axis variable
interaction.plot( 
    df$Sanitation,#x-axis variable
    df$Smoking,#variable for line
    df$Lifetime) #y-axis variable
```

#Measure of Accuracy
ROC Curve and AUC value
```{r}
library(ROCit)
## Warning: package 'ROCit' was built under R version 3.5.2
ROCit_obj <- rocit(mod$fitted.values,df$Life_Expectancy)
plot(ROCit_obj)
ROCit_obj$AUC
```

Confusion Matrix
```{r}
predict_life <- predict(mod,df,"response")
predict_life <- as.factor(ifelse(predict_life > 0.5, "Above Average", "Below Average"))
confusionMatrix(predict_life,df$Life_Expectancy,mode = "everything")

```

Using CV to test for overfitting
```{r}
set.seed(827)

#Train test split; using 80% as training
sample <- sample(c(TRUE, FALSE), nrow(df), replace=TRUE, prob=c(0.8,0.2))
train  <- df[sample, ]
test   <- df[!sample, ]

# Cross-Validation
set.seed(123)
train.control <- trainControl(method = "cv", 
                              number = 10, repeats = 3)
# Train the model
model <- train(Life_Expectancy~log(Consumption)+Education+Depression+Smoking+Sanitation+Smoking*Education, train, 
               method = "glm",family="binomial",
               trControl = train.control)

model$resample
```



#Interpretation
With confidence interval of exponentiated coefficients, we can interpret the Odds and Logit Scores
```{r}
c1 = exp(coef(mod))
c2 = exp(confint(mod))
cbind(Estimate=c1,c2)

```

Plot of Odds
```{r}
library(sjPlot)
plot_model(mod)
```

```{r,warning=F}
mmps(mod,~log(Consumption)+Education+Diet_Plant+Obesity+Depression+Smoking+Sanitation+ Smoking*Sanitation)
```
```{r}
influencePlot(mod)
```


#Variance Analysis
```{r}
#better residual plot by binning into categories
library(arm)
plot(mod,1)
binnedplot(fitted(mod), 
           residuals(mod, type = "response"), 
           nclass = NULL, 
           xlab = "Expected Values", 
           ylab = "Average residual", 
           main = "Binned residual plot", 
           cex.pts = 0.8, 
           col.pts = 1, 
           col.int = "gray")
```


